---
title: "Homework 05: Factor and Figure Management"
author: "Aidan Hughes"
date: "October 17, 2018"
output: 
  github_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

Taken from the [homework HTML file](http://stat545.com/Classroom/assignments/hw05/hw05.html), here are the goals for this assignment:


- Reorder a factor in a principled way based on the data and demonstrate the effect in arranged data and in figures.

- Write some data to file and load it back into R.

- Improve a figure (or make one from scratch), using new knowledge, e.g., control the color scheme, use factor levels, smoother mechanics.

- Make a plotly visual.

- Implement visualization design principles.

# Part 1: Factor Management

First things first, load all of our packages.

```{r packages}
suppressPackageStartupMessages(library("gapminder"))
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("singer"))
suppressPackageStartupMessages(library("scales"))
```

## Elaborating Gapminder Dataset

Before manipulating the levels of gapminder, let's first check the columns are actually factors and then look at some numbers so we can do a before/after comparison.

```{r before dropping}
nrow(gapminder)

cont <- gapminder$continent

is.factor(cont)
nlevels(cont)
levels(cont)
```

```{r after dropped}
no_Oceania <- gapminder %>%
  filter(continent != "Oceania") %>%
  droplevels()
  
nlevels(no_Oceania$continent)
levels(no_Oceania$continent)
nrow(no_Oceania)
```

We can plot the before and after data frames to see the difference. 

```{r Oceania not dropped}
gapminder %>%
  mutate(continent = fct_reorder(continent, gdpPercap, .fun = min)) %>%
  ggplot(aes(continent, gdpPercap)) +
  geom_jitter(colour = "darkblue", alpha = 0.3) +
  scale_y_log10(labels=comma_format()) + 
  labs(y = "GDP Per Capita", x = "Continent", title = "GDP Per Capita of Continents Before Dropping Oceania") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_minimal()
```


```{r Oceania dropped}
no_Oceania %>%
  mutate(continent = fct_reorder(continent, gdpPercap, .fun = min)) %>%
  ggplot(aes(continent, gdpPercap)) +
  geom_jitter(colour = "orange", alpha = 0.3) +
  scale_y_log10(labels=comma_format()) + 
  labs(y = "GDP Per Capita", x = "Continent", title = "GDP Per Capita of Continents After Dropping Oceania") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_minimal()
```

Clearly, the GDP per Capita no longer plots Oceania as it has been dropped.


We can try `arrange()` as well with gdpPercap, which would be the same as using `fct_reorder` with the default`mean` function.

```{r no Oceania plot}
no_Oceania %>%
  arrange(gdpPercap) %>%
  ggplot(aes(continent, gdpPercap)) +
  geom_jitter(colour = "darkgreen", alpha = 0.3) +
  scale_y_log10(labels=comma_format()) + 
  labs(y = "GDP Per Capita", x = "Continent", title = "GDP Per Capita of Continents Arranged by Mean") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_minimal()

```

## Elaborating Singer Dataset

Let's have a quick look at the structure of the `singer_locations` dataset.

```{r singer dataset}
head(singer_locations)
nrow(singer_locations)
```

We can see that none of the columns are currently factors, so let's convert the year column into a factor before dropping any data.

```{r factor year}
# Using the base R function
singerFactorYear1 <- singer_locations %>%
  mutate(year = as.factor(year))

# Using the forcats function
singerFactorYear2 <- singer_locations %>%
  mutate(year = factor(year))

tibble("as.factor()" = singerFactorYear1$year,
       "factor()" = singerFactorYear2$year) %>%
  head() %>%
  knitr::kable()

```


In this case, the output of the base R factor function and the forcats function are the same, but its alway's good to check in case of surprises.

Now we can move onto dropping the "0" years.

```{r drop 0 years}
singerFixedYears <- singerFactorYear1 %>%
  filter(year != 0) %>%
  mutate(year = fct_drop(year))

nrow(singerFixedYears)

any(levels(singerFixedYears) == 0)
```

Remembering from before, we originally had 10100 entries in the `singer_locations` dataset, and now the number has decreased. As well, the unused level "0" has been dropped.

Now, let's try rearranging `artist_name` to show which songs had the highest "hotness" score.

```{r}
singer_locations %>%
  mutate(artist_name = factor(artist_name)) %>%
  mutate(artist_name = fct_reorder(artist_name, artist_hotttnesss, .fun = max)) %>%
  select(artist_name, artist_hotttnesss, year) %>%
  head() %>%
  knitr::kable()
```

Good job, Motion City Sountrack.

# Part 2: File I/O