---
title: "Homework 05: Factor and Figure Management"
author: "Aidan Hughes"
date: "October 17, 2018"
output: 
  github_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

Taken from the [homework HTML file](http://stat545.com/Classroom/assignments/hw05/hw05.html), here are the goals for this assignment:


- Reorder a factor in a principled way based on the data and demonstrate the effect in arranged data and in figures.

- Write some data to file and load it back into R.

- Improve a figure (or make one from scratch), using new knowledge, e.g., control the color scheme, use factor levels, smoother mechanics.

- Make a plotly visual.

- Implement visualization design principles.

# Part 1: Factor Management

First things first, load all of our packages.

```{r packages}
suppressPackageStartupMessages(library("gapminder"))
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("singer"))
```

Let's manipulate some factors in gapminder.

```{r}
gap_2002 <- gapminder %>%
  select(country) %>%
  distinct(country) %>%
  filter(country %in% c("Canada", "Mexico"))
  
countries <- gap_2002$country

class(countries)
nlevels(countries)
countries
```



